Vragenlijst thesis evaluatie

### 2015-06-17 doc/template/chap-9.tex
Kleine veranderingen aan diff flip en flipvirus

\subsubsection{Single resource}
The basic FlipIt game consists of a single resource. To represent the security problem of the propagation of an APT in a network, the adapted game defines its single resource as a computer network with multiple
nodes. One of the players, the defender, will also try to defend his network. The defender
will do this by flipping all the nodes of the network (i.e. the entire resource) in every move he plays. The
attacker on the other hand will try to infect all the nodes in the network. For an attacker on the other hand, it is quite impossible to infect all computers in a network at once. Rather, the attacker will try to identify the node in the graph that can infect all the nodes in the
shortest possible time. Then, after dropping a virus on this first node, it takes a while for the virus to infect the entire network. 
%Since the original FlipIt game works with a single resource that is always flipped entirely, the assumption is made that the attacker is considered to gain immediate full control over the resource when the network has been infected, even it is only one node that has been infected.\\
\subsubsection{Only one action per player}

The action of the defender consist of defending his network. In this context it will be defending every computer controlled device in a network. To make the network clean the defender can patch the systems with the latest security bulletins or reinstall the software or even wipe out the computer. All these different actions can imply other costs. The cost of patching a system will be smaller than the cost of wiping out a computer. \\
The defender has only one action. Sending different viruses and worms to a computer network. Every time the attacker flips the resource, he does this by sending a new virus to the system. The attacker does not send the same virus again, because if the computers are patched, the virus cannot get in any more. 

@@

@@

Moves in the basic FlipIt game are stealthy or covert and not immediately detected by the other payer. The attacker attacker's moves are stealthy because we want to model a scenario where a computer network is attacker by an APT. The main characteristic of an APT is that the attack is stealthy. Depending on how the attacker has set up his attack, and depending on the network if it is connected with the Internet or not, the moves of the defender are stealthy or not stealthy. For example, if the attacker launches an APT only to harm the system and not to receive feedback, the moves of the defender are stealthy. If the attacker launches an APT to steal sensitive information, the moves of the defender are non stealthy because the attacker can see that he doesnt receive information anymore. But the defender can also setup a honey pot, making the attacker believe that he is stealing sensitive information, but the information in the honeypot is all fake. In thesis we assume that both moves of the attacker and the defender are stealthy. -- kunnen enerzijds niet stealthy zijn als the apt informatie steelt depending how he setup his attacker. depending on network connected. in this a thesis assumed, non stealthy is..

@@

@@

\subsubsection{Motivation Application}
\subsubsection*{Wormlike ATP}

>>>>

\subsubsection{Single resource}
The basic FlipIt game consists of a single resource. To represent the security problem of the propagation of an APT in a network, the adapted game defines its single resource as a computer network with multiple
nodes. One of the players, the defender, will also try to defend his network.
% The defender will do this by flipping all the nodes of the network (i.e. the entire resource) in every move he plays. 
The defender will do this by flipping the nodes of the network.
The attacker on the other hand will try to infect all the nodes in the network. 
%For an attacker on the other hand, it is quite impossible to infect all computers in a network at once. Rather, t
The attacker will try to identify the node in the graph that can infect all the nodes in the shortest possible time. Then, after dropping a virus on this first node, it takes a while for the virus to infect the entire network. 
%Since the original FlipIt game works with a single resource that is always flipped entirely, the assumption is made that the attacker is considered to gain immediate full control over the resource when the network has been infected, even it is only one node that has been infected.\\
\subsubsection{Only one action per player}

The action of the defender consist of defending his network. The network consists of multiple nodes. The defender flips all the nodes of the network in every move he plays. The action of the defender can be extended to the action where the defender only flips a subset of the nodes of the network. This extention has already been investigated in the context of placing anti-virus systems in a graph. This problem is known to be NP-hard. [\todo{referentie}].
The defender has only one action. Sending different viruses and worms to a computer network. Every time the attacker flips the resource, he does this by sending a new virus to the system. The attacker does not send the same virus again, because if the computers are patched, the virus cannot get in any more. Sending a virus in this context is equal to flipping a node in the network. The node can be a targeted node or a random node. If the attacker has knowledge about the graph, the attacker can choose to target the node that can infect all the nodes in the shortest time possible.

@@

The moves of the defender are immediate. It is assumed that if a defender tries to clean the network that this will happen without a delay. The defender can de-plug the computer from the network, or clean. 

@@

Moves in the basic FlipIt game are stealthy or covert and not immediately detected by the other payer. The attacker attacker's moves are stealthy because we want to model a scenario where a computer network is attacker by an APT. The main characteristic of an APT is that the attack is stealthy. Depending on how the attacker has set up his attack, and depending on the network if it is connected with the Internet or not, the moves of the defender are stealthy or not stealthy. For example, if the attacker launches an APT only to harm the system and not to receive feedback, the moves of the defender are stealthy. If the attacker launches an APT to steal sensitive information, the moves of the defender are non stealthy because the attacker can see that he doesn not receive information any more. But the defender can also set-up a honey pot, making the attacker believe that he is stealing sensitive information, but the information in the honeypot is all fake. In thesis we assume that both moves of the attacker and the defender are stealthy. -- kunnen enerzijds niet stealthy zijn als the apt informatie steelt depending how he setup his attacker. depending on network connected. in this a thesis assumed, non stealthy is..

@@

 The defender will be defending every computer controlled device in a network. To make the network clean the defender can patch the systems with the latest security bulletins or reinstall the software or even wipe out the computer. All these different actions can imply other costs. The cost of patching a system will be smaller than the cost of wiping out a computer. \\ 
\todo{cost of the attacker ook toevoegen?}

@@

\subsubsection{Motivation Application: Wormlike ATP}
